\mytitle{}

### 2019-07-27 paper/infocom.tex
Work-in-progress trimming/rewording/new stuff...

@@

\section{Introduction}
?? Redo intro to be more system focussed? Needs to be suitably different.

@@

%?? Discuss the most important conclusions before the outline.
We find that the existing work for this task fails to account for congestion-aware traffic (i.e., TCP) and deployment environments with high host density per egress point, achieving poor results due to an overly coarse view of the network.
The approach we identify to remedy this is to make throttling decisions on a per-source basis, and we present the engineering decisions this mandates: updating RL agents from multiple traces per timestep, timed random sequential action computation and a supporting \emph{software-defined network} (SDN) architecture.
In tandem with the development and evaluation of an effective state space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
Our introduction of per-source decisions improves substantially upon the state-of-the-art when acting upon most internet traffic (i.e., congestion-aware protocols), and we show that our second model achieves excellent performance for high host density in this case.
Crucially, both models remain protocol- and content-agnostic to offer future-proofing against the rollout of future protocols like QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}.
%?? Also algorithmic enhancements such as multiple actions per timestep, 
%?? PROTOCOL-AGNOSTIC -- HOW WILL THESE THINGS COPE WITH QUIC ET AL.?!?!
Briefly, this paper contributes:
\begin{itemize}
%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
%	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}. We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
	\item Two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
	\item Reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
	\item An empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation using these new traffic models (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
\end{itemize}

\section{Background and Threat Model}
%?? Introduce RL, related definitions etc.

@@

\emph{Distributed denial of service} (DDoS) attacks are concentrated efforts by many hosts to reduce the availability of a service, typically to inflict financial harm or as an act of vandalism.
Attackers achieve this by either exploiting peculiarities of operating system or application behaviour in \emph{semantic attacks} (e.g., \emph{SYN flooding attacks}), or overwhelming their target through sheer volume of requests or inbound packets (\emph{volume-based attacks}) \cite{DBLP:conf/imc/JonkerKKRSD17}.
Hosts often participate unwillingly, typically having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.

Although there are variations of each class of attack, flooding attacks are the most relevant to our work.
\emph{Amplification attacks} exploit the presence of services who eagerly send large replies in response to small requests, where UDP-based services like DNS and NTP are most exploitable \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
Malicious hosts send many small requests, spoofed to appear as though they originated from the victim, causing many large replies to be sent to the intended target---significantly increasing a botnet's throughput while masking the identity of each participant.
\emph{Transit-link/link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.

\subsection{Reinforcement Learning}\label{sec:reinforcement-learning}
\emph{Reinforcement learning} (RL) is a variant of machine learning concerned with training an agent to choose an optimal sequence of actions for a given task \cite{RL2E}.
We assume at any point in time $t$, an agent knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
This model of system interaction is best described as a \emph{Markov Decision Process} (MDP).
RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken.
The policy is refined as more decisions and rewards become available, learning adaptively and online.
In practice, this means that agents are able to automatically adapt to evolving problems without operator intervention or a new training corpus.

@@

Tile coding is a form of feature representation which converts a state-action pair into a sparse boolean feature vector $\operatorname{\mathbf{x}}(s, a)$ by subdividing a $d$-dimensional subset of the space into a number of overlapping grids.
Each tile corresponds to an entry of $\operatorname{\mathbf{x}}(s, a)$ which is set to 1 if the state-action pair lies within it.
\Cref{fig:tilecode-code} demonstrates the process for a 2-dimensional state space, and that the numbers of tilings and tiles per dimension control feature resolution and generalisation.
We may then approximate an action's value with respect to a policy parameter vector $\wvec{}$, defining some $\acval{s}{a}{\wvec{}} \approx \operatorname{q}(s, a)$:
\begin{equation}
%	\begin{gather}
\acval{s}{a}{\wvec{}} = \wvec{}^{\top} \operatorname{\mathbf{x}}(s, a)
%	\end{gather}
\label{eqn:lin-approx}
\end{equation}
As each component of $\wvec{}$ is the value estimate of the corresponding tile, learning an effective policy is equivalent to learning $\wvec{}$.
Given a learning rate $\alpha \in \mathbb{R}$ and initialising $\wvec{0}=\bm{0}$, we may continually update $\wvec{t}$ using the \emph{1-step semi-gradient Sarsa} algorithm \cite[pp.\ \numrange{243}{244}]{RL2E}:
\begin{subequations}
	\begin{gather}
	\delta_t = R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}} - \acval{S_t}{A_t}{\wvec{t}},\\
	\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla{\acval{S_t}{A_t}{\wvec{t}}},
	\end{gather}%
	\label{eqn:sg-sarsa}%
	where $\delta_t$ is known as the \emph{temporal-difference} (TD) error, the discount factor $\gamma \in [0,1)$ determines how crucial future rewards are, and the vector gradient $\nabla$ is taken with respect to $\wvec{}$.
\end{subequations}

@@

\subsection{Motivation}\label{sec:motivation}
%?? What makes RL a suitable method for network anomaly detection, what features are most relevant?
%?? Point I was thinking of: feedback-loop-like model allows monitoring \emph{after} an action is taken to (in theory) allow forgiveness of mistakenly punished flows. This does hinge on taking a flow-by-flow look at the state space, but if we can combine knowledge of current state (duh!), the last action taken (i.e. an indicator of our previous assessment [such as high pdrop $\implies$ bad]) then perhaps a flow which falls off identically to a legit flow can be rescued.
Moving beyond the overt benefits of choosing RL-based defences for coping with non-stationary problems, we believe that there are concrete reasons for their use here.
We have seen that for other domains in particular, misclassification is a serious problem, which can introduce \emph{collateral damage} in the context of DDoS prevention.
In theory, the feedback-loop-like model allows us to monitor flows \emph{after} an action is taken to allow forgiveness of mistakenly punished flows.
This does rely upon the ability to take a flow-by-flow view of the state space, but if we can combine knowledge of current state with the last applied action, then perhaps a flow which falls off identically to a legitimate flow can be rescued.

%Which features might be best suited to this problem?
%?? Relevant features: aggregate network state (load at various points [this has been done, of course]), flow-specific measurements (upload/download ratio when bandwidth above threshold \cite{DBLP:conf/ndss/Rossow14}, packet inter-arrival times, etc.)
Other studies suggest that there are particularly useful features which make the task of online DDoS flow identification feasible.
Aggregate network load observed at various locations suggests the overall health of a network \cite{DBLP:journals/eaai/MalialisK15}, and the ratio of correspondence between pair flows can suggest asymmetry and in many cases illegitimacy \cite{DBLP:conf/ndss/Rossow14}.
Generic volume-based statistics (counts, counts per duration, average packet sizes) have seen effectiveness in such as $k$-nearest neighbours classifiers trained to detect DDoS attacks in progress \cite{DBLP:conf/dsn/LeeKSPY17}.
Most importantly, there is evidence showing behavioural changes in response to bandwidth expansion \cite{DBLP:conf/ndss/KangGS16}, suggesting similar artefacts might arise after throttling, packet drop, or other interference.
%?? If we assume amplification attacks, we know it won't be `random' source IPs (since it's mostly-legit servers who think that they're doing a good job by replying)
%?? If we assume amplification attacks, we know it won't be `random' source IPs (since it's mostly-legit servers who think that they're doing a good job by replying)

@@

\subsection{Threat Model}
An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to any host which connects to it, and they are expected to act rationally in its pursuit.
Threat actors are external and act intentionally, aren't expected to be classified as an \emph{advanced persistent threat}, and likely range from hacktivists to moderately funded adversaries.
We assume that attacks will be volumetric DDoS attacks targeting bandwidth with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
The addresses of the set of unwitting reflector nodes are visible to the target, though any bots taking part in an attack or the machines those bots control are not revealed to the target without communication/collaboration with \nth{3} party organisations such as upstream ISPs.
The discovery of any reflector by some defence system does have a cost to the attacker---there is a particularly large (yet finite) supply of viable reflector nodes \cite{DBLP:conf/ndss/Rossow14}, but the constraints that each has a large upstream bandwidth and support for high-amplification-factor protocols narrow this pool.

We do not assume that an attacker has white-box access to the parameters underlying an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
While they may be able to perform some degree of reverse engineering by observing the health of their own legitimate canary flows, ``stealing'' the policy through observation \cite{DBLP:conf/uss/TramerZJRR16}, investigating whether perturbations would persist in a medium as volatile as network traffic statistics falls outside of the scope of this work.
The same observation extends to the possibility of poisoning attacks \cite{DBLP:journals/jmlr/KloftL10, DBLP:conf/acsac/ShenTS16}; to the best of our knowledge, no studies have been undertaken surrounding the mis-training of RL agents.
These are APT-level capabilities, whose exploration presents a rich source for future work.

\section{DDoS Mitigation with Per-flow Reinforcement Learning}\label{sec:ddos-mitigation-with-per-flow-reinforcement-learning}
Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to modify the design of agents such that filtering decisions are computed per-flow.
However, any of these alterations must account for computational constraints imposed by the deployment environment---the amount of flows passing over an agent is unbounded.
We describe and justify our approach, our algorithmic improvements, and present two action models, one of which draws on domain knowledge introduced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.

\subsection{System Design and Assumptions}
A deployment environment is defined as a network, with a set of \emph{ingress/egress points} from its domain of control, through which traffic can flow, and a set of protected \emph{destination nodes}.
These destination nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
Agents are co-located with each egress switch, and control the proportion of upstream packets from each external host to discard.
Each destination node $s$ has a maximum capacity, $U_s$.

We assume that the deployment environment is a moderately complex software-defined network, because the paradigm offers features which can directly benefit RL agents acting within.
The OpenFlow protocol allows a controller (or other authorised hosts) to install complex actions, forwarding rules and logic into a switch at runtime.
Furthermore, networks of this kind more naturally enable the future use of \emph{network function virtualisation}, a technology which could allow relocation and easy installation of learners.
Agents communicate with their co-hosted OpenFlow-enabled switches---running a modified version of \emph{Open vSwitch} (OVS) \cite{open-vswitch}---to install probabilistic packet-drop rules.

\subsection{Algorithm}
To make decisions cheaply and at low latency, we use \emph{semi-gradient Sarsa with tile coding} as described in \cref{eqn:sg-sarsa} and \cref{sec:reinforcement-learning}, rather than making use of neural networks or more complicated function approximators.
Exploration is introduced by making use of $\epsilon$-greedy action selection, linearly annealing $\epsilon$ to 0 over time.
Each agent has its own internal parameter vector $\wvec{}$, and agents do not share their weight vector updates with one another (but may share experience and traces with one another).

Although the choice of a classical RL method likely brings lower theoretical performance, there are significant reasons to favour such methods; these include lower latency decision-making, lower energy usage, reduced model complexity (and training time), the availability of necessary hardware, and simpler decision boundaries.
This aligns with our goal of quick online learning, and potentially faster adaptation to aggregate changes in traffic without introducing dedicated tensor processing hardware to networks.
Simpler decision boundaries reduce the risk of overfitting and unexpected behaviour, and we anticipate that the simplicity of policy computation in the chosen manner will also greatly aid interpretability of anomalous action choices.

When choosing a learning algorithm, we compared against Q-learning as well as methods based on \emph{eligibility traces} such as Watkins's $\operatorname{Q}(\lambda)$ \cite[pp. 312--314]{RL2E} and $\operatorname{Sarsa}(\lambda)$ \cite[pp. 305]{RL2E}.
Preliminary experiments found that Sarsa offered the best performance and behaviour.
?? Link plot or table w Sarsa vs. Q results.

@@

		\end{tikzpicture}
	}
	\caption{
		Global state selection for a flow between an external host and server $s_0$ which passes over Agent 1.
		All nodes in the path taken through the defended network are filled in blue, and all link load measurements which are chosen for action computation are indicated with a thick blue line.
		\label{fig:global-state-path}
	}
\end{figure}

@@

\subsubsection{Risks}
Our mode of action means that each agent is in control of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}, and so carries a risk of introducing collateral damage into the network.
This is particularly severe when handling TCP traffic: the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97} states that TCP bandwidth is proportional to $1/\sqrt{p}$ (noting that $p$ is nonzero in any real network) while constant bitrate (CBR) UDP traffic is proportional to $1 - p$.
%It's worth noting that there are various ways that this could be implemented, and that the application of \emph{programmable data planes} to this end are suggested as future work.
This weakness still presents itself in modern flavours of TCP, such as TCP Cubic which in turn has bandwidth proportional to $1/p^{0.75}$ \cite{rfc8312}.
This is of particular importance due to the prevalence of TCP and other congestion-aware protocols within the Internet.
Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data volume\footnote{Repository link removed for double-blind reviewing.}.
%Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data volume\footnote{\url{https://github.com/FelixMcFelix/caida-stats}}.
QUIC, a future congestion-aware protocol, comprises \SIrange{2.6}{9.1}{\percent} of traffic observed on backbone links, depending on location and typical workload \cite{DBLP:conf/pam/RuthPDH18}.
%As far as future network protocols are concerned, QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}, a congestion-aware stream transmission protocol, will behave much like TCP, showing the importance of further development to properly handle traffic with such characteristics.

%?? Make more sane. This is about ``Why go per-flow?''
This further justifies our focus on per-flow decisions---real-world deployments see many flows pass over any egress point, making global actions (such as those chosen by \textcite{DBLP:journals/eaai/MalialisK15}) more likely to inflict collateral damage.
%This manifests in two ways: the best-achievable performance drops, and so too does the learning rate.
Given the probability that a host is legitimate, $P_G \in [0,1]$,  if it is assumed that a host may be only legitimate or malicious then it follows that it will be malicious with probability $P_B = 1 - P_G$.
Supposing we have a set of $n$ hosts, the probability that all hosts will be legitimate is $P_G^n$ (resp. $P_B$ for malicious).
Defining \emph{imperfect service} $P_{M,n}$ to mean any case where all $n$ hosts connecting over a switch do not share the same classification (i.e., a mixture), then the probability that a switch is delivering imperfect service is $P_{M,n} = 1 - (P_G^n + P_B^n)$.
\begin{thm}
	As the host/learner ratio $n$ increases, it is more likely that a throttling switch will exhibit imperfect service: $\forall n \in \mathbb{Z}^{+}, P_{M,n} \le P_{M,n+1}$.
\end{thm}
\begin{proof}
	\emph{Base case:} $P_{M,1}=0, P_{M,2} = 1 - P_G^2 - P_B^2 \ge 0$.

	\emph{Inductive step:} Assume that the theorem holds for $n$. Observe that $P_G^n \ge P_G^{n+1}$ (resp.\ $P_B$). It then follows that:
	\begin{align*}
	P_G^n + P_B^n &\ge P_G^{n+1} + P_B^{n+1}\\
	1 - (P_G^n + P_B^n) &\le 1 - (P_G^{n+1} + P_B^{n+1})\\
	P_{M,n} &\le P_{M,n+1} \qedhere
	\end{align*}
\end{proof}
\begin{corr}
	Restricting $P_G \in (0,1)$ so that both $P_G$ and $P_B$ are non-zero ensures strict inequality: $P_{M,n} < P_{M,n+1}$.
\end{corr}
When considering that many hosts have an especially adverse reaction to our main means of control, flow-level granularity becomes an obvious choice.

@@

%?? Here is where I would put my ramblings about system architecture...
We present our design of a system which operates and support the effective real-world deployment of RL-based agents, acting to mitigate DDoS attacks.
\Cref{fig:sys-arch} displays this visually, separating system elements which are local to each agent from those which reside elsewhere in the network.
We describe the main purpose and operation of each module within an agent's \emph{virtual function} (VF), and discuss techniques to make deployment more efficient using existing technologies.

\subsection{Core and RL Executor}
The core module is the main loop for decision-making in an agent's architecture.
At each timestep, the core receives information about which flows have arrived and should be acted upon from the \emph{TRS Scheduler}.
The core then retrieves the current and previous state vector associated with each flow from the \emph{Flowstate Database}, passing them into the RL algorithm alongside the last action chosen for that flow (if available).

>>>>

\usepackage{etoolbox}
\expandafter\patchcmd\csname citeauthor \endcsname
{\begingroup}{\begingroup\aftergroup\@}{}{}

@@

\section{Introduction}
?? Redo intro to be more system focussed? Needs to be suitably different. Perhaps reduce focus on evolving, make it way more about "protocol-agnostic"...

@@

%?? Discuss the most important conclusions before the outline.
We find that the existing work for this task fails to account for congestion-aware traffic (i.e., TCP) and deployment environments with high host density per egress point, achieving poor results due to an overly coarse view of the network.
The approach we identify to remedy this is to make throttling decisions on a per-source basis, and we present the engineering decisions this mandates: updating RL agents from multiple traces per timestep, timed random sequential action computation and a supporting \emph{software-defined network} (SDN) architecture.
In tandem with the development and evaluation of an effective state space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
Our introduction of per-source decisions improves substantially upon the state-of-the-art when acting upon most internet traffic (i.e., congestion-aware protocols), and we show that our second model achieves excellent performance for high host density in this case.
Crucially, both models remain protocol- and content-agnostic to offer future-proofing against the rollout of future protocols like QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}.
%?? Also algorithmic enhancements such as multiple actions per timestep, 
%?? PROTOCOL-AGNOSTIC -- HOW WILL THESE THINGS COPE WITH QUIC ET AL.?!?!
Briefly, this paper contributes:
\begin{itemize}
%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
%	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}. We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
	\item Two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
	\item A description of the system architecture which supports the real-world deployment of such agents (\cref{sec:system-architecture}).
	\item An investigation into suitable features for an automatic DDoS mitigation system (\cref{sec:rethinking-the-state-space}).
%	\item Reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
	\item A comparison of the effectiveness of different RL algorithms in these agents (\cref{sec:the-results-of-doing-so}).
	\item An empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
\end{itemize}

\section{Background}
%?? Introduce RL, related definitions etc.

@@

\emph{Distributed denial of service} (DDoS) attacks are concentrated efforts by many hosts to reduce the availability of a service, typically to inflict financial harm or as an act of vandalism.
Attackers achieve this by exploiting operating system or application behaviour in \emph{semantic attacks} (e.g., \emph{SYN flooding attacks}), or by overwhelming their target through volume of requests or traffic (\emph{volume-based attacks}) \cite{DBLP:conf/imc/JonkerKKRSD17}.
Hosts often participate unwillingly, typically having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.

Volume-based attacks are the most relevant to our work, with two varieties being especially prevalent today.
\emph{Amplification attacks} exploit services who send large replies in response to small requests, with UDP-based services like DNS and NTP seeing most abuse \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
Malicious hosts send many small requests, spoofed using the victim's address, causing many large replies to be sent to the intended target---increasing a botnet's throughput while leaving each participant anonymous.
\emph{Transit-link/link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.

\subsection{Reinforcement Learning}\label{sec:reinforcement-learning}
\emph{Reinforcement learning} (RL) is a variant of machine learning concerned with training an agent to choose an optimal sequence of actions for a given task \cite{RL2E}.
We assume at any point in time $t$, an agent knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
This model of system interaction is best described as a \emph{Markov Decision Process} (MDP), where agents aim to maximise the \emph{expected reward} received.
RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken.
The policy is refined as more decisions and rewards become available, learning online.
In practice, this means that agents are able to automatically adapt to evolving problems without operator intervention or a new training corpus.

@@

Tile coding is a form of feature representation which converts a state-action pair into a sparse boolean feature vector $\operatorname{\mathbf{x}}(s, a)$ by subdividing a $d$-dimensional subset of the space into a number of overlapping grids.
Each tile corresponds to an entry of $\operatorname{\mathbf{x}}(s, a)$ which is set to 1 if the state-action pair lies within it.
\Cref{fig:tilecode-code} demonstrates the process for a 2-dimensional state space, and that the numbers of tilings and tiles per dimension control feature resolution and generalisation.
We may then approximate an action's value with respect to a policy parameter vector $\wvec{}$, defining some $\acval{s}{a}{\wvec{}} \approx \operatorname{q}(s, a)$:
\begin{equation}
%	\begin{gather}
\acval{s}{a}{\wvec{}} = \wvec{}^{\top} \operatorname{\mathbf{x}}(s, a)
%	\end{gather}
\label{eqn:lin-approx}
\end{equation}
As each component of $\wvec{}$ is the value estimate of the corresponding tile, learning an effective policy is equivalent to learning $\wvec{}$.
Given a learning rate $\alpha \in \mathbb{R}$ and initialising $\wvec{0}=\bm{0}$, we may continually update $\wvec{t}$ using the \emph{1-step semi-gradient Sarsa} algorithm \cite[pp.\ \numrange{243}{244}]{RL2E}:
\begin{subequations}
	\begin{gather}
	\delta_t = R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}} - \acval{S_t}{A_t}{\wvec{t}},\\
	\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla{\acval{S_t}{A_t}{\wvec{t}}},
	\end{gather}%
	\label{eqn:sg-sarsa}%
	where $\delta_t$ is known as the \emph{temporal-difference} (TD) error, the \emph{discount factor} $\gamma \in [0,1)$ determines how crucial future rewards are, and the vector gradient $\nabla$ is taken with respect to $\wvec{}$.
\end{subequations}

@@

\subsection{Motivation}\label{sec:motivation}
%?? What makes RL a suitable method for network anomaly detection, what features are most relevant?
%?? Point I was thinking of: feedback-loop-like model allows monitoring \emph{after} an action is taken to (in theory) allow forgiveness of mistakenly punished flows. This does hinge on taking a flow-by-flow look at the state space, but if we can combine knowledge of current state (duh!), the last action taken (i.e. an indicator of our previous assessment [such as high pdrop $\implies$ bad]) then perhaps a flow which falls off identically to a legit flow can be rescued.
There are concrete reasons to use RL methods in this domain, beyond handling non-stationarity.
Misclassification can be a serious problem, introducing \emph{collateral damage} in a DDoS prevention system.
RL's feedback-loop model lets us monitor flows \emph{after} an action is taken to perhap allow forgiveness of mistakenly punished flows.
This relies upon a flow-by-flow view of the state space, but by combining knowledge of current state with the last applied action, a punished flow which falls off identically to a legitimate flow can be rescued.

%Which features might be best suited to this problem?
%?? Relevant features: aggregate network state (load at various points [this has been done, of course]), flow-specific measurements (upload/download ratio when bandwidth above threshold \cite{DBLP:conf/ndss/Rossow14}, packet inter-arrival times, etc.)
Other studies suggest that there are particularly useful features which make the task of online DDoS flow identification feasible.
Aggregate network load observed at various locations suggests the overall health of a network \cite{DBLP:journals/eaai/MalialisK15}, and the ratio of correspondence between pair flows can suggest illegitimacy \cite{DBLP:conf/ndss/Rossow14}.
Volume-based statistics (counts, counts per duration, average packet sizes) have seen use in $k$-nearest neighbours classifiers trained to detect DDoS attacks in progress \cite{DBLP:conf/dsn/LeeKSPY17}.
Most importantly, there is evidence showing that bandwidth expansion may cause behavioural changes  which differ between benign and attack traffic \cite{DBLP:conf/ndss/KangGS16}.
%?? If we assume amplification attacks, we know it won't be `random' source IPs (since it's mostly-legit servers who think that they're doing a good job by replying)
%?? If we assume amplification attacks, we know it won't be `random' source IPs (since it's mostly-legit servers who think that they're doing a good job by replying)

@@

%\subsection{Threat Model}
%An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to any host which connects to it, and they are expected to act rationally in its pursuit.
%Threat actors are external and act intentionally, aren't expected to be classified as an \emph{advanced persistent threat}, and likely range from hacktivists to moderately funded adversaries.
%We assume that attacks will be volumetric DDoS attacks targeting bandwidth with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
%The addresses of the set of unwitting reflector nodes are visible to the target, though any bots taking part in an attack or the machines those bots control are not revealed to the target without communication/collaboration with \nth{3} party organisations such as upstream ISPs.
%The discovery of any reflector by some defence system does have a cost to the attacker---there is a particularly large (yet finite) supply of viable reflector nodes \cite{DBLP:conf/ndss/Rossow14}, but the constraints that each has a large upstream bandwidth and support for high-amplification-factor protocols narrow this pool.
%
%We do not assume that an attacker has white-box access to the parameters underlying an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
%While they may be able to perform some degree of reverse engineering by observing the health of their own legitimate canary flows, ``stealing'' the policy through observation \cite{DBLP:conf/uss/TramerZJRR16}, investigating whether perturbations would persist in a medium as volatile as network traffic statistics falls outside of the scope of this work.
%The same observation extends to the possibility of poisoning attacks \cite{DBLP:journals/jmlr/KloftL10, DBLP:conf/acsac/ShenTS16}; to the best of our knowledge, no studies have been undertaken surrounding the mis-training of RL agents.
%These are APT-level capabilities, whose exploration presents a rich source for future work.

\section{DDoS Mitigation with Per-flow Reinforcement Learning}\label{sec:ddos-mitigation-with-per-flow-reinforcement-learning}
%Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to modify the design of agents such that filtering decisions are computed per-flow.
We hypothesise that per-flow agent designs will offer better accuracy than current RL-based agents for DDoS mitigation.
%However, any of these alterations must account for computational constraints imposed by the deployment environment---the amount of flows passing over an agent is unbounded.
We describe and justify our approach, our algorithmic improvements, and present two action models, one of which draws on domain knowledge introduced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.

\subsection{System Design and Assumptions}
A deployment environment is a network, with a set of \emph{ingress/egress points} from its domain of control through which traffic can flow, and a set of protected \emph{destination nodes}.
These destination nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
Agents are co-located with each egress switch, and control the proportion of upstream packets from each external host to discard.
Each destination node $s$ has a maximum capacity, $U_s$.

\subsection{Algorithm}
To make decisions and learn cheaply with low latency, we use \emph{semi-gradient Sarsa with tile coding} \ref{eqn:sg-sarsa} (\cref{sec:reinforcement-learning}), rather than make use of neural networks.
Exploration is introduced by making use of $\epsilon$-greedy action selection, linearly annealing $\epsilon$ to 0 over time.
Each agent has its own internal parameter vector $\wvec{}$.

While classical RL methods likely have lower accuracy, there are reasons to favour them; faster decisions, lower energy usage, reduced model complexity (and training time), and the availability of compatible hardware.
This aligns with our goal of quick online learning, and potentially faster adaptation to aggregate changes in traffic without introducing dedicated tensor processing hardware to networks.
Approaches like this have simpler decision boundaries, reducing the risk of overfitting and unexpected behaviour. 
Simpler methods of computing policies may also aid in interpreting anomalous action choices.

@@

		\end{tikzpicture}
	}
	\caption{
		Global state selection for a flow between an external host and server $s_0$ which passes over Agent 1.
		Nodes in the path taken are filled in blue, and link load measurements which are used for decisions are indicated with a thick blue line.
		\label{fig:global-state-path}
	}
\end{figure}

@@

\subsubsection{Risks}
Our mode of action means that each agent is in control of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}, and so carries a risk of introducing collateral damage into the network.
This is particularly severe when handling TCP traffic: the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97} states that TCP bandwidth is proportional to $1/\sqrt{p}$ (noting that $p$ is nonzero in any real network) while constant bitrate (CBR) UDP traffic is proportional to $1 - p$.
%It's worth noting that there are various ways that this could be implemented, and that the application of \emph{programmable data planes} to this end are suggested as future work.
This weakness still presents itself in modern flavours of TCP, such as TCP Cubic which in turn has bandwidth proportional to $1/p^{0.75}$ \cite{rfc8312}.
This is of particular importance due to the prevalence of TCP and other congestion-aware protocols within the Internet.
Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data volume\footnote{Repository link removed for double-blind reviewing.}.
%Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data volume\footnote{\url{https://github.com/FelixMcFelix/caida-stats}}.
QUIC, a future congestion-aware protocol, comprises \SIrange{2.6}{9.1}{\percent} of traffic observed on backbone links, depending on location and typical workload \cite{DBLP:conf/pam/RuthPDH18}.
Furthermore, applying a single rule to many flows (as in \textcite{DBLP:journals/eaai/MalialisK15}) increases the likelihood of mistaken collateral damage.
When considering that many hosts have an especially adverse reaction to our main means of control, flow-level granularity becomes an obvious choice.

@@

%?? Here is where I would put my ramblings about system architecture...
We present our design of a system which operates and support the effective real-world deployment of RL-based agents, acting to mitigate DDoS attacks.
\Cref{fig:sys-arch} displays this visually, separating system elements which are local to each agent from those which reside elsewhere in the network.
In particular, we assume that each agent operates as a \emph{virtual function} (VF) adjacent to a software-defined switch.
Software-defined networks allow a controller (or other authorised hosts) to install complex actions, forwarding rules and logic into a switch at runtime.
Furthermore, networks of this kind more naturally enable the future relocation and easy installation of learners.
Agents communicate with their co-hosted OpenFlow-enabled switches---running a modified version of \emph{Open vSwitch} (OVS) \cite{open-vswitch}---to install probabilistic packet-drop rules.
We describe the main purpose and operation of each module within an agent's VF, and discuss techniques to make deployment more efficient using existing technologies.
